Comparing Optimization Algorithms in the Fitting of Linear Mixed Models: Evaluating Speed and Accuracy using \pkg{lme4} in R and \pkg{lmm} in Julia

### 2015-08-03 inst/doc/Paper.Rnw
Edited results section, tried to fix issues with latex table.  Note that pdf from rstuio will not directly compile as written, but pdf file is still created correctly.

\begin{verbatim}
  | Row | opt        | dsname          | n      | np | excess  | time    | reltime|    
  |-----|------------|-----------------|--------|----|---------|---------|-------|
  | 1   | "LD\_CCSAQ" | "Alfalfa"       | 72     | 1  | 0.0     | 0.0017  | 27.5171|
  | 2   | "LD\_CCSAQ" | "AvgDailyGain"  | 32     | 1  | 0.0     | 0.0014  | 13.9236|
  | 3   | "LD\_CCSAQ" | "AvgDailyGain"  | 32     | 1  | 0.0     | 0.0014  | 10.8009|
  | 4   | "LD\_CCSAQ" | "BIB"           | 24     | 1  | 0.0     | 0.0013  | 11.7005|
  | 5   | "LD\_CCSAQ" | "Bond"          | 21     | 1  | 0.0     | 0.0009  | 13.6969|
  | 6   | "LD\_CCSAQ" | "bs10"          | 1104   | 20 | 0.0     | 1.0958  | 12.608|
  | 7   | "LD\_CCSAQ" | "bs10"          | 1104   | 8  | 39.9948 | 0.0375  | 23.2699|
  | 8   | "LD\_CCSAQ" | "cake"          | 270    | 1  | 0.0     | 0.0033  | 18.8753|
  | 9   | "LD\_CCSAQ" | "Cultivation"   | 24     | 1  | 0.0     | 0.0009  | 15.6477|
  | 10  | "LD\_CCSAQ" | "Demand"        | 77     | 2  | 3.21928 | 0.0055  | 22.0827|
  | 11  | "LD\_CCSAQ" | "dialectNL"     | 225866 | 6  | 0.0     | 6.9896  | 3.942|
  .
  .
  .
  | 19  | "LD\_CCSAQ" | "gb12"          | 512    | 8  | 103.176 | 0.0218  | 13.1358|
  | 20  | "LD\_CCSAQ" | "HR"            | 120    | 3  | 0.0     | 0.0089  | 8.6481|
  | 21  | "LD\_CCSAQ" | "Hsb82"         | 7185   | 1  | 192.73  | 0.0102  | 3.7438|
  | 22  | "LD\_CCSAQ" | "IncBlk"        | 24     | 1  | 0.55726 | 0.001   | 5.0422|
  | 23  | "LD\_CCSAQ" | "kb07"          | 1790   | 72 | 8.20739 | 17.4698 | 9.2732|
  | 24  | "LD\_CCSAQ" | "Mississippi"   | 37     | 1  | 0.93471 | 0.0006  | 15.9582|
  | 25  | "LD\_CCSAQ" | "mm0"           | 69588  | 6  | 0.0     | 4.8286  | 1.1034|
  | 26  | "LD\_CCSAQ" | "Oxboys"        | 234    | 3  | 136.788 | 0.0169  | 5.3039|
  | 27  | "LD\_CCSAQ" | "PBIB"          | 60     | 1  | 0.0     | 0.0014  | 8.6177|
  | 28  | "LD\_CCSAQ" | "Penicillin"    | 144    | 2  | 0.0     | 0.0131  | 15.3763|
  | 29  | "LD\_CCSAQ" | "Semiconductor" | 48     | 1  | 0.0     | 0.0012  | 1.133|
  | 30  | "LD\_CCSAQ" | "SIMS"          | 3691   | 3  | 3.60856 | 0.134   | 2.3154|

@@

>>>>

\begin{verbatim}
  | Row | opt        | dsname          | n      | np | excess  | time    | reltime|    
  |-----|------------|-----------------|--------|----|---------|---------|-------|
  | 1   | "LD\_CCSAQ" | "Alfalfa"       | 72     | 1  | 0.0     | 0.0017  | 27.5171|
  | 2   | "LD\_CCSAQ" | "AvgDailyGain"  | 32     | 1  | 0.0     | 0.0014  | 13.9236|
  | 3   | "LD\_CCSAQ" | "AvgDailyGain"  | 32     | 1  | 0.0     | 0.0014  | 10.8009|
  | 4   | "LD\_CCSAQ" | "BIB"           | 24     | 1  | 0.0     | 0.0013  | 11.7005|
  | 5   | "LD\_CCSAQ" | "Bond"          | 21     | 1  | 0.0     | 0.0009  | 13.6969|
  | 6   | "LD\_CCSAQ" | "bs10"          | 1104   | 20 | 0.0     | 1.0958  | 12.608|
  | 7   | "LD\_CCSAQ" | "bs10"          | 1104   | 8  | 39.9948 | 0.0375  | 23.2699|
  | 8   | "LD\_CCSAQ" | "cake"          | 270    | 1  | 0.0     | 0.0033  | 18.8753|
  | 9   | "LD\_CCSAQ" | "Cultivation"   | 24     | 1  | 0.0     | 0.0009  | 15.6477|
  | 10  | "LD\_CCSAQ" | "Demand"        | 77     | 2  | 3.21928 | 0.0055  | 22.0827|
  | 11  | "LD\_CCSAQ" | "dialectNL"     | 225866 | 6  | 0.0     | 6.9896  | 3.942|
  | 12  | "LD\_CCSAQ" | "Dyestuff2"     | 30     | 1  | 0.0     | 0.0005  | 0.682|
  | 13  | "LD\_CCSAQ" | "Dyestuff"      | 30     | 1  | 0.0     | 0.0008  | 1.0184|
  | 14  | "LD\_CCSAQ" | "ergoStool"     | 36     | 1  | 0.0     | 0.0012  | 1.1018|
  | 15  | "LD\_CCSAQ" | "Exam"          | 4059   | 1  | 0.0     | 0.0105  | 1.1851|
  | 16  | "LD\_CCSAQ" | "Exam"          | 4059   | 1  | 0.0     | 0.0105  | 1.318|
  | 17  | "LD\_CCSAQ" | "Gasoline"      | 32     | 1  | 0.0     | 0.0011  | 0.8193|
  | 18  | "LD\_CCSAQ" | "gb12"          | 225866 | 6  | 0.0     | 3.5221  | 17.4791|
  | 19  | "LD\_CCSAQ" | "gb12"          | 512    | 8  | 103.176 | 0.0218  | 13.1358|
  | 20  | "LD\_CCSAQ" | "HR"            | 120    | 3  | 0.0     | 0.0089  | 8.6481|
  | 21  | "LD\_CCSAQ" | "Hsb82"         | 7185   | 1  | 192.73  | 0.0102  | 3.7438|
  | 22  | "LD\_CCSAQ" | "IncBlk"        | 24     | 1  | 0.55726 | 0.001   | 5.0422|
  | 23  | "LD\_CCSAQ" | "kb07"          | 1790   | 72 | 8.20739 | 17.4698 | 9.2732|
  | 24  | "LD\_CCSAQ" | "Mississippi"   | 37     | 1  | 0.93471 | 0.0006  | 15.9582|
  | 25  | "LD\_CCSAQ" | "mm0"           | 69588  | 6  | 0.0     | 4.8286  | 1.1034|
  | 26  | "LD\_CCSAQ" | "Oxboys"        | 234    | 3  | 136.788 | 0.0169  | 5.3039|
  | 27  | "LD\_CCSAQ" | "PBIB"          | 60     | 1  | 0.0     | 0.0014  | 8.6177|
  | 28  | "LD\_CCSAQ" | "Penicillin"    | 144    | 2  | 0.0     | 0.0131  | 15.3763|
  | 29  | "LD\_CCSAQ" | "Semiconductor" | 48     | 1  | 0.0     | 0.0012  | 1.133|
  | 30  | "LD\_CCSAQ" | "SIMS"          | 3691   | 3  | 3.60856 | 0.134   | 2.3154|

@@

The \texttt{time} column is the time in seconds to converge. The
\texttt{reltime} column is the time relative to the \texttt{LN\_BOBYQA}
optimizer in the \pkg{MixedModels} package for \proglang{Julia}.

For \proglang{Julia} the time column is the time in seconds to converge. The reltime column is the time relative to the LN_BOBYQA optimizer in the \pkg{MixedModels} package for \proglang{Julia}.

%%% Note that because of this table, rstudio cannot compile pdf because of margins,
% but pdf is still available (and compiled properly) as seperate file.
\begin{verbatim}
| Row | opt               | dsname          | n      | np | excess | time   | reltime|
|-----|-------------------|-----------------|--------|----|--------|--------| ----|
| 1|"NLOPT\_LN\_BOBYQA" | "Alfalfa"      | 72     | 1  | 0.0    | 0.042  | 27.5171|
| 2|"NLOPT\_LN\_BOBYQA" | "Animal"       | 20     | 2  | 0.0    | 0.023  |13.9236|
| 3|"NLOPT\_LN\_BOBYQA" | "Assay"        | 60     | 2  | 1.0e-5 | 0.032  |10.8009|
| 4|"NLOPT\_LN\_BOBYQA" | "AvgDailyGain" | 32     | 1  | 0.0    | 0.02   |11.7005|
| 5|"NLOPT\_LN\_BOBYQA" | "AvgDailyGain" | 32     | 1  | 0.0    | 0.02   |13.6969|
| 6|"NLOPT\_LN\_BOBYQA" | "BIB"          | 24     | 1  | 0.0    | 0.02   |12.608|
| 7|"NLOPT\_LN\_BOBYQA" | "Bond"         | 21     | 1  | 0.0    | 0.02   |23.2699
| 8|"NLOPT\_LN\_BOBYQA" | "bs10"         | 1104   | 20 | 1.0e-5 | 4.661  |18.8753|
| 9|"NLOPT\_LN\_BOBYQA" | "bs10"         | 1104   | 8  | 0.0    | 1.057  |15.6477|
| 10|"NLOPT\_LN\_BOBYQA" | "cake"        | 270    | 1  | 0.0    | 0.053  |22.0827
| 11|"NLOPT\_LN\_BOBYQA" | "Chem97"      | 31022  | 2  | 0.0    | 0.632  |3.942|
.
.
.
| 38|"NLOPT\_LN\_BOBYQA" | "PBIB"        | 60     | 1  | 0.0    | 0.018  |13.1358|
| 39|"NLOPT\_LN\_BOBYQA" | "Penicillin"  | 144    | 2  | 0.0    | 0.023  |8.6481|
| 40|"NLOPT\_LN\_BOBYQA" | "Poems"       | 275996 | 3  | 0.0    | 21.309 |3.7438|
| 41|"NLOPT\_LN\_BOBYQA" | "ScotsSec"    | 3435   | 2  | 0.0    | 0.076  |5.0422|
| 42|"NLOPT\_LN\_BOBYQA" | "Semi2"       | 72     | 3  | 0.0    | 0.03   |9.2732|
| 43|"NLOPT\_LN\_BOBYQA" | "Semiconductor"| 48     | 1  | 0.0    | 0.019  |15.9582|
| 44|"NLOPT\_LN\_BOBYQA" | "SIMS"        | 3691   | 3  | 0.0    | 0.15   |1.1034|
| 45|"NLOPT\_LN\_BOBYQA" | "sleepstudy"  | 180    | 3  | 0.0    | 0.037  |5.3039|
| 46|"NLOPT\_LN\_BOBYQA" | "sleepstudy"  | 180    | 2  | 0.0    | 0.024  |15.3763|
| 47|"NLOPT\_LN\_BOBYQA" | "TeachingII"  | 96     | 1  | 0.0    | 0.021  |1.133|
| 48|"NLOPT\_LN\_BOBYQA" | "Weights"     | 399    | 3  | 1.0e-5 | 0.039  |5.3039|
| 49|"NLOPT\_LN\_BOBYQA" | "WWheat"      | 60     | 3  | 0.0    | 0.025  |2.3154|
\end{verbatim}

### 2015-08-03 inst/doc/Paper.tex
Edited results section, tried to fix issues with latex table.  Note that pdf from rstuio will not directly compile as written, but pdf file is still created correctly.

>>>>

The \texttt{time} column is the time in seconds to converge. The
\texttt{reltime} column is the time relative to the \texttt{LN\_BOBYQA}
optimizer in the \pkg{MixedModels} package for \proglang{Julia}.

For \proglang{Julia} the time column is the time in seconds to converge. The reltime column is the time relative to the LN_BOBYQA optimizer in the \pkg{MixedModels} package for \proglang{Julia}.

\begin{verbatim}
| Row | opt               | dsname          | n      | np | excess | time   | reltime|
|-----|-------------------|-----------------|--------|----|--------|--------| ----|
| 1|"NLOPT\_LN\_BOBYQA" | "Alfalfa"      | 72     | 1  | 0.0    | 0.042  | 27.5171|
| 2|"NLOPT\_LN\_BOBYQA" | "Animal"       | 20     | 2  | 0.0    | 0.023  |13.9236|
| 3|"NLOPT\_LN\_BOBYQA" | "Assay"        | 60     | 2  | 1.0e-5 | 0.032  |10.8009|
| 4|"NLOPT\_LN\_BOBYQA" | "AvgDailyGain" | 32     | 1  | 0.0    | 0.02   |11.7005|
| 5|"NLOPT\_LN\_BOBYQA" | "AvgDailyGain" | 32     | 1  | 0.0    | 0.02   |13.6969|
| 6|"NLOPT\_LN\_BOBYQA" | "BIB"          | 24     | 1  | 0.0    | 0.02   |12.608|
| 7|"NLOPT\_LN\_BOBYQA" | "Bond"         | 21     | 1  | 0.0    | 0.02   |23.2699
| 8|"NLOPT\_LN\_BOBYQA" | "bs10"         | 1104   | 20 | 1.0e-5 | 4.661  |18.8753|
| 9|"NLOPT\_LN\_BOBYQA" | "bs10"         | 1104   | 8  | 0.0    | 1.057  |15.6477|
| 10|"NLOPT\_LN\_BOBYQA" | "cake"        | 270    | 1  | 0.0    | 0.053  |22.0827
| 11|"NLOPT\_LN\_BOBYQA" | "Chem97"      | 31022  | 2  | 0.0    | 0.632  |3.942|
.
.
.
| 38|"NLOPT\_LN\_BOBYQA" | "PBIB"        | 60     | 1  | 0.0    | 0.018  |13.1358|
| 39|"NLOPT\_LN\_BOBYQA" | "Penicillin"  | 144    | 2  | 0.0    | 0.023  |8.6481|
| 40|"NLOPT\_LN\_BOBYQA" | "Poems"       | 275996 | 3  | 0.0    | 21.309 |3.7438|
| 41|"NLOPT\_LN\_BOBYQA" | "ScotsSec"    | 3435   | 2  | 0.0    | 0.076  |5.0422|
| 42|"NLOPT\_LN\_BOBYQA" | "Semi2"       | 72     | 3  | 0.0    | 0.03   |9.2732|
| 43|"NLOPT\_LN\_BOBYQA" | "Semiconductor"| 48     | 1  | 0.0    | 0.019  |15.9582|
| 44|"NLOPT\_LN\_BOBYQA" | "SIMS"        | 3691   | 3  | 0.0    | 0.15   |1.1034|
| 45|"NLOPT\_LN\_BOBYQA" | "sleepstudy"  | 180    | 3  | 0.0    | 0.037  |5.3039|
| 46|"NLOPT\_LN\_BOBYQA" | "sleepstudy"  | 180    | 2  | 0.0    | 0.024  |15.3763|
| 47|"NLOPT\_LN\_BOBYQA" | "TeachingII"  | 96     | 1  | 0.0    | 0.021  |1.133|
| 48|"NLOPT\_LN\_BOBYQA" | "Weights"     | 399    | 3  | 1.0e-5 | 0.039  |5.3039|
| 49|"NLOPT\_LN\_BOBYQA" | "WWheat"      | 60     | 3  | 0.0    | 0.025  |2.3154|
\end{verbatim}
